Evolving Motion Machines

### 2019-04-26 Final/paper.tex
...The rest of the paper. I apparently forgot to commit progress made on results last night.

\subsection{Evaluation System}

\section{Results}

\section{Discussion}

\section{Future Work}

%different distance measurement (relative angle)
%different terrain types
%different evaluation point (center of mass)
%cyclic joints in agent structure.
\section{Conclusion}

>>>>

\subsection{Evaluation System}
A proper goal function to optimize is critical to any evolutionary system.
Agents in this system are given the goal of traveling as far from their initial position as possible in each generation.
Generations are bounded by a generation length parameter that determines how much time agents are given within a generation to produce movement.
Four agents are run in parallel, set not to collide with each other, and the agent that travels the greatest distance in that time is used to seed the next generation.
Initially, agents consist only of their root atom and a single random child.
Any further variations are caused by the evolutionary system.

Determining distance is a non-trivial task however.
This system interprets distance as the amount of distance travelled by the root atom of that agent.
While all the pieces of an agent can move in space, only changes in the root atom are recorded.
This calculation is performed by summing the change in global position of the root atom in every frame of the physics simulation.
This includes motion in all three axes, and thus includes motion types such as spinning, shaking, and rolling.

\section{Results}
\begin{figure}
    \begin{center}
    \includegraphics[width=1\textwidth]{DistanceChart}
    \end{center}
    \caption[ ]{
        A scatter plot showing both the best and average distances in meters achieved in each generation, as well as a trend line for each metric.
        The variance in the best distance attained is a product of Unity's physics engine being non-deterministic.
    } 
    \label{fig:DistanceChart} 
\end{figure}

Before fully discussing the specific results of the system, two major caveats need to be discussed.
First is that the system, by design, has an incredibly large search space.
The intention was to create a system that has significant variance in the agents it produces.
The consequence, however, is that finding the global optimum is nearly computationally impossible.

This issue is further exacerbated by the following, more difficult problem.
This experimentation has shown that Unity's physics engine is not deterministic, which creates difficulties in evaluating the performance of an agent.
Using the default settings provided in the simulator, significant variance can be observed in the outcome of even simple physics tests such as dropping an object on the ground.
For large, complex agents, this variance expands exponentially and can cause an effective agent to fail when re-evaluated.
This behaviour was not noticed until observing the thrashing of best results visible in Fig. 1.

To address this issue, the simulation's settings were increased to best performance.
This somewhat alleviated the issue, but results still varied.
The results shown in Fig. 1. use the optimized settings, but the thrashing is still very apparent.
This also incurred a significant performance penalty, such that even slightly outdated computers may experience slow down.

However, even with this thrashing behaviour occuring the system provides positive results.
Distance covered trends upwards at a fairly linear rate throughout the system's lifetime.
Most improvements make simple point mutations to a small number of atoms, slightly increasing the agent's overall performance.
Others cause significant vertical jumps in the agent's performance and are usually caused by a significant change in structure.
Overall, despite the thrashing that is pervasive in the results, agents do successfully grow both in structure and performance.

The actual structure generated by agents tends to fall into to main categories.
Agents in the first category have 2-4 long, mostly static branches which jut out from the agent's root atom. 
These branches swing around wildly, causing the agent to jump and bounce in a mostly random fashion.
In the second category, agents make large flat spinning discs.
These discs are imbalanced, such that the whole rolls on the edge of the disc approximately 20 degrees off the ground.
Most of the time, agents in the second category look like off-balance wheels, while the first look like flailing arms and legs.

These results are very strong indicators that this system is working as intended.
This system does seem to favour larger, more complex agents rather than simple agents that use a simple motion type.
Further, the abstract shape of agents that emerges is surprisingly rational very frequently.

\section{Future Work}
While this system is a very positive result, there are many ways in which the system can be improved.
First and foremost would be to increase the number of motion types available.
Atoms can currently either rotate or not move at all, but a linear thrusting type of motion may increase the number of different types of agents.
Another approach would be to allow cyclic joints in the agents structure, creating larger, stronger structures from the base atoms.
Finally, by using the agent's center of mass to measure distance, the thrashing behaviour may be addressed further.

There are also other potential extensions of the system thtat might be of benefit that do not require changing the agent's structure.
First would be to modify how distance is measured.
Many agents in the current system abuse the cyclic motion types to spin in place or roll in small, tight circles.
To address this, a distance measurement that penalizes angled movements may create different overall behaviours.
By penalizing angles, rather than jumping and spinning agents would likely learn to move in continuous straight lines.

Another approach would be to introduce varied terrain, rather than the current terrain that is simply a large, flat plane.
Contrary to the approach above, introducing hills and valleys may cause vastly different behaviour.
It's difficult to determine whether the same structures generated for this terrain would work for terrains.

\section{Conclusion}
This system has shown strong results while taking a less common approach to evolutionary systems.
The two research goals, creating several diverse solutions and creating complex moving structures, have both been met.
However, the system is still clearly in an infantile state, and there is lots of room for improvement.
This is yet another positive for this system, as further development is likely to produce more and more interesting results.
